Clustering time series based on probability distributions across temporal granularities

### 2021-11-17 main.Rmd
cluster summaries

Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, and \ref{fig:parcoord-application}. In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups. --> Groups are colored differently whenever they represent different groupings. While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.

@@

```{r combined-groups-js, fig.cap = "The distribution of electricity demand across groups over hod (a), moy (b), and hod (conditional on wnwd) (c) are shown through quantile plots. Group 1 (1-3), Group 2 (4-9), Group 3 (10 - 15), Group 4 (16- 20), Group 5 (21 - 24). Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year). With their conditional dependencies on hod or moy categories, wnwd differences may not be discernible on their own. (c) demonstrates that the distribution of hod conditional on wnwd is also not different, except for the tails. This might be because most of these chosen customers do not have a wnwd pattern. It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics. To facilitate this comparison, the individual distributions are sorted by their JS-based groupings.", eval = FALSE}

@@

```{r groups-4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown. Groups A-1, A-2, and A-3 profiles correspond to Groups B-1, B-2, and B-3, respectively. A-4 is split into B-4 and B-5, each of which has a distinct shape across moy and wnwd. B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage. B-5 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}

>>>>

Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:combined-groups-js} and \ref{fig:parcoord-application}. (XXX move the next sentences down to where you talk about the clustering results for JS) In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups. --> Groups are colored differently whenever they represent different groupings. While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.

@@

```{r combined-groups-js, fig.cap = "The distribution of electricity demand across clusters over $hod$ (a), $moy$ (b), and $wnwd$ (c) are shown through quantile plots. Group 1 (customers 1-3), Group 2 (customers 4-9), Group 3 (customers 10-15), Group 4 (customers 16-20), Group 5 (customers 21-24). Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year). With their conditional dependencies on hod or moy categories, wnwd differences may not be discernible on their own. (c) demonstrates that the distribution of hod conditional on wnwd is also not different, except for the tails. This might be because most of these chosen customers do not have a wnwd pattern. It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics. To facilitate this comparison, the individual distributions are sorted by their JS-based groupings.", eval = FALSE}

@@

```{r groups-4and5, fig.cap = "Summary plots for four and five clusters. The distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$ is shown. Groups A-1 (customers XXX), A-2 (customers XXX), and A-3 (customers XXX) profiles correspond to Groups B-1, B-2, and B-3, respectively. Cluster A-4 splits into B-4 (customers XXX) and B-5 (customers XXX)to produce the five clusters, whci h better resolves the $moy$ distribution. B-4 is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage. B-5 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range. When $k = 4$ is used, these two groups merge to form A-4, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=5}
